% we include the glossary here (frontmatter is included with \input, so this command is as if it was in main.tex)
\input{frontmatter/glossary}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\maketitlepage


%----------------------------------------------------------------------------------------
%	STATEMENT of INTEGRITY
%----------------------------------------------------------------------------------------
\integritystatement

%----------------------------------------------------------------------------------------
%	DEDICATION  (optional)
%----------------------------------------------------------------------------------------
%
%\dedicatory{For/Dedicated to/To my\ldots}
% \begin{dedicatory}
% The dedicatory is optional. Below is an example of a humorous dedication.
%
% "To my wife Marganit and my children Ella Rose and Daniel Adam without whom this book would have been completed two years earlier." in "An Introduction To Algebraic Topology" by Joseph J. Rotman.
% \end{dedicatory}

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}

% here you put the abstract in the main language of the work.

O presente trabalho de dissertação propõe o desenvolvimento do \gls{FUSE}, uma plataforma de software inovadora concebida para centralizar e organizar sistemas de videovigilância heterogéneos, geograficamente dispersos e situados em redes não controladas. Esta investigação surge em resposta a desafios críticos identificados na segurança pública e privada, nomeadamente a fragmentação tecnológica, as vulnerabilidades de segurança inerentes às comunicações \gls{P2P} e a ineficiência operacional enfrentada por entidades como os \glspl{OPC} na análise forense de vídeo.

A arquitetura proposta integra uma camada de abstração de hardware que normaliza a comunicação com câmaras de diversos fabricantes através do protocolo \gls{RTSP}, garantindo a interoperabilidade. A segurança das comunicações é assegurada através de túneis \gls{VPN} que protegem a integridade e confidencialidade dos dados em redes não controladas, complementada por um modelo de controlo de acessos \gls{ABAC}. Adicionalmente, o sistema incorpora um módulo de visão computacional progressivo, estruturado em três fases: deteção de atividade para filtragem temporal, classificação de objetos (e.g., humanos, veículos) e extração de atributos específicos, visando a automatização da análise forense e a redução significativa da carga de trabalho manual.

A metodologia adotada segue o paradigma de "Design and Creation", com a validação da solução a ser realizada através de um  \gls{MVP} em cenários simulados e reais. Os resultados esperados centram-se na demonstração da eficácia da integração segura de dispositivos, na robustez da proteção de dados e na otimização dos processos de investigação criminal, sempre em estrita observância dos princípios éticos e regulamentares, como o EU AI Act.


\end{abstract}

\begin{abstractotherlanguage}
% here you put the abstract in the "other language": English, if the work is written in Portuguese; Portuguese, if the work is written in English.

This dissertation proposes the development of \gls{FUSE}, an innovative software platform designed to centralize and organize heterogeneous, geographically dispersed video surveillance systems located in uncontrolled networks. This research emerges in response to critical challenges identified in public and private security, namely technological fragmentation, security vulnerabilities inherent to \gls{P2P} communications, and the operational inefficiency faced by entities such as \glspl{LEA} in forensic video analysis.

The proposed architecture integrates a hardware abstraction layer that normalizes communication with cameras from various manufacturers via the \gls{RTSP} protocol, ensuring interoperability. Communication security is ensured through \gls{VPN} tunnels that protect data integrity and confidentiality across uncontrolled networks, complemented by an \gls{ABAC} model. Additionally, the system incorporates a progressive computer vision module structured in three phases: activity detection for temporal filtering, object classification (e.g., humans, vehicles), and specific attribute extraction, aiming to automate forensic analysis and significantly reduce manual workload.

The adopted methodology follows the "Design and Creation" paradigm, with the solution's validation being conducted through an \gls{MVP} in both simulated and real-world scenarios. The expected results focus on demonstrating the effectiveness of secure device integration, the robustness of data protection, and the optimization of criminal investigation processes, always in strict compliance with ethical and regulatory principles such as the EU AI Act.

\end{abstractotherlanguage}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS (optional)
%----------------------------------------------------------------------------------------

\begin{acknowledgements}

Agradeço ao meu orientador, Professor Paulo Baltarejo Sousa, pela disponibilidade, orientação e apoio prestado ao longo do desenvolvimento deste trabalho. Agradeço igualmente à entidade acolhedora, StabilityBubble, Lda., por disponibilizar os recursos e os meios profissionais que tornaram o desenvolvimento do projeto mais acessível, bem como pelo enquadramento e pelas ligações com o contexto real que contribuíram para aproximar a componente teórica da prática.
\end{acknowledgements}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents

\listoffigures % Prints the list of figures

\listoftables % Prints the list of tables

% \iflanguage{portuguese}{
% \renewcommand{\listalgorithmname}{Lista de Algor\'itmos}
% }
% \listofalgorithms % Prints the list of algorithms
% \addchaptertocentry{\listalgorithmname}


% \renewcommand{\lstlistlistingname}{List of Source Code}
% \iflanguage{portuguese}{
% \renewcommand{\lstlistlistingname}{Lista de C\'odigo}
% }
% \lstlistoflistings % Prints the list of listings (programming language source code)
% \addchaptertocentry{\lstlistlistingname}


%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------
%\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)
%%\textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere\\
%%\textbf{WSF} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or\\
%\end{abbreviations}

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

%\begin{symbols}{lll} % Include a list of Symbols (a three column table)
%
%$a$ & distance & \si{\meter} \\
%$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
%%Symbol & Name & Unit \\
%
%\addlinespace % Gap to separate the Roman symbols from the Greek
%
%$\omega$ & angular frequency & \si{\radian} \\
%
%\end{symbols}



%----------------------------------------------------------------------------------------
%	ACRONYMS
%----------------------------------------------------------------------------------------

\newcommand{\listacronymname}{List of Acronyms}
\iflanguage{portuguese}{
\renewcommand{\listacronymname}{Lista de Acr\'onimos}
}

%Use GLS
\glsresetall
\printglossary[title=\listacronymname,type=\acronymtype,style=long]

%----------------------------------------------------------------------------------------
%	DONE
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering
\pagestyle{thesis} % Return the page headers back to the "thesis" style
